---
title: "R Notebook"
output: html_notebook
---

```{r, Warnings = FALSE}
library(tidyverse)
library(dplyr)
library(mosaic)
library(ggplot2)
```

##Variables: 

PEAK ID- This is a unique identifier for the mountain
Name - Name of the climber
Yr/Seas- Year and Season of the ascent
Date- Date of Ascent
Time- Time of Ascent
Citizenship-
Sex
Age
Oxygen - Ascent with/without Oxygen
Dth- Did the ascent result in death?
Host - Country

##Sample Datasets in the order that we created them 
- StratifiedSample
```{r}
#Import data
myFile <- "https://raw.githubusercontent.com/shravanisamala/Mount_Everest_Project/master/Mt_Everest_Ascent_Data.csv"

MountEverestData <- read.csv(myFile)
#MountEverestData <- read.csv("C:\\Users\\shrav\\Desktop\\Stat 466\\Mount_Everest_Project\\Mt_Everest_Ascent_Data.csv")
colnames(MountEverestData)[1] <- gsub('^...','',colnames(MountEverestData)[1])

head(MountEverestData)

# #Cleaning up data
# MountEverestData <- 
#   MountEverestData %>%
#   rename(Peak = Peak., 
#          Name = Name., 
#          Yr.Seas = Yr.Seas., 
#          Date = Date., 
#          Time = Time., 
#          Citizenship = Citizenship., 
#          Sex = Sex., 
#          Age = Age., 
#          Oxy = Oxy., 
#          Dth = Dth., 
#          Host = Host.)
  
MountEverestData$Dth <- as.factor(ifelse(MountEverestData$Dth == "Y", 1, 0)) #1 = death, 0 = alive
MountEverestData$Oxy <- as.factor(ifelse(MountEverestData$Oxy == "Y", 1, 0)) #1 = death, 0 = alive
MountEverestData$YearOnly <- as.integer(substr(MountEverestData$Yr.Seas, 1, 4))
MountEverestData$Helper <- as.integer(substr(MountEverestData$Yr.Seas, 1, 3))


MountEverestData <- 
MountEverestData %>% mutate(Decade =
                     case_when(Helper == 195 ~ "1950",
                               Helper == 196 ~ "1960",
                               Helper == 197 ~ "1970",
                               Helper == 198 ~ "1980",
                               Helper == 199 ~ "1990",
                               Helper == 200 ~ "2000",
                               Helper == 201 ~ "2010",
                               Helper == 202 ~ "2020")
)


MountEverestData$Helper <- NULL # removes Helper column because its irrelavent lol 
#View(MountEverestData)
head(MountEverestData, 10)
```

### 1. Our population of interest are the all of the adults that climbed Mount Everest from 1953 through 1920

### 2. One parameter of interest is the population age. The population mean age is 34.69089. 
```{r}
mu = mean(MountEverestData$Age)
mu

avg = sum(MountEverestData$Age) / length(MountEverestData$Age)
avg
```
### 3. Sample Size for Stratified Random Sampling 
- As we can see Yr.Seas and Citizenship have way too many possible strata 
- Host , Oxy, Death, and Sex seem to be homogenoue like we need. 
- We made Decade a variable and that looks like it has a lot of variability 
```{r}
#Trying to find Strata
ggplot(
  data = MountEverestData, 
  aes(x = Host, y = Age)) +
  geom_boxplot()  + 
  labs(title = "Host vs. Age")

ggplot(
  data = MountEverestData, 
  aes(x = Yr.Seas, y = Age)) +
  geom_boxplot()  + 
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Year vs. Age")

ggplot(
  data = MountEverestData, 
  aes(x = Citizenship, y = Age)) +
  geom_boxplot()  + 
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Citizenship vs. Age")

ggplot(
  data = MountEverestData, 
  aes(x = Sex, y = Age)) +
  geom_boxplot()  + 
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Sex vs. Age")

ggplot(
  data = MountEverestData, 
  aes(x = Oxy, y = Age)) +
  geom_boxplot()  + 
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Oxygen vs. Age")

ggplot(
  data = MountEverestData, 
  aes(x = Dth, y = Age)) +
  geom_boxplot()  + 
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Death vs. Age")

ggplot(
  data = MountEverestData, 
  aes(x = YearOnly, y = Age)) +
  geom_boxplot()  + 
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Year Only vs. Age")

ggplot(
  data = MountEverestData, 
  aes(x = Decade, y = Age)) +
  geom_boxplot()  + 
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Decade vs. Age")


```

Step 1: n = (sum (N_k * sd_k))^2 / N^2D + sum(N_i * sd_i)^2
```{r}
# Find N

Stratified <- 
  MountEverestData %>%
  group_by(Sex)%>%
  summarise(n(), sd(Age), mean(Age))
Stratified <- 
  Stratified %>%
  rename(count = "n()", 
         sd = "sd(Age)", 
         mean = "mean(Age)")
Stratified['count_Age'] <- Stratified[2]
Stratified['sd_Age'] <- Stratified[3]
Stratified['mean_Age'] <- Stratified[4]


Stratified$count <- NULL
Stratified$sd <- NULL
Stratified$mean <- NULL
Stratified

```

We will use Neyman Allocation. So N = 10184, and we will make our Bound 3. So n = 47
```{r}

x = ((Stratified$count_Age[1] * Stratified$sd_Age[1]) + (Stratified$count_Age[2] * Stratified$sd_Age[2]))^2

N = length(MountEverestData$Age)
B = 3
D = (B^2) / 4

y = ((Stratified$count_Age[1] * Stratified$sd_Age[1]^2) + (Stratified$count_Age[2] * Stratified$sd_Age[2]^2))
n = x/((N^2 * D) + y)
n
```
We must sample 3 females and 44 males. 
```{r}
nFemale = 47* ((Stratified$count_Age[1] * Stratified$sd_Age[1]) / ((Stratified$count_Age[1] * Stratified$sd_Age[1]) + (Stratified$count_Age[2] * Stratified$sd_Age[2])))

nMale = 47* ((Stratified$count_Age[2] * Stratified$sd_Age[2]) / ((Stratified$count_Age[1] * Stratified$sd_Age[1]) + (Stratified$count_Age[2] * Stratified$sd_Age[2])))

nFemale
nMale
```

4. Yes, the strata seem to be homogenous. Their means and variances look very similar.
```{r}
#a. 
#Creating first Dataset
MountEverestDataFem <- 
  MountEverestData %>%
  filter(Sex == "F")

MountEverestDataMale <- 
  MountEverestData %>%
  filter(Sex == "M")

StratifiedSample1 <- sample(MountEverestDataFem, 3, replace = FALSE)
StratifiedSample2 <- sample(MountEverestDataMale, 44, replace = FALSE)

StratifiedSample <- rbind(StratifiedSample1, StratifiedSample2)
StratifiedSample

#b. 
ggplot(
  data = MountEverestData, 
  aes(x = Sex, y = Age)) +
  geom_boxplot()  + 
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Sex vs. Age")
```
ybarst = (1/N) * sum(NiYi)
```{r}
#c 
### This is the final dataset that we use for calculations. However, it will change every time we sample / run previous code
StratifiedSampleFinal <- 
  StratifiedSample %>%
    group_by(Sex) %>%
    summarise(n(), sd(Age), mean(Age)) %>%
  rename(count = "n()", 
         sd = "sd(Age)", 
         mean = "mean(Age)")

StratifiedSampleFinal['count_Age_Sample'] <- StratifiedSampleFinal[2]
StratifiedSampleFinal['sd_Age_Sample'] <- StratifiedSampleFinal[3]
StratifiedSampleFinal['mean_Age_Sample'] <- StratifiedSampleFinal[4]


StratifiedSampleFinal$count <- NULL
StratifiedSampleFinal$sd <- NULL
StratifiedSampleFinal$mean <- NULL
StratifiedSampleFinal

y_bar_st = (1/N) * ((Stratified$count_Age[1] * StratifiedSampleFinal$mean_Age_Sample[1]) + (Stratified$count_Age[2] * StratifiedSampleFinal$mean_Age_Sample[2]))
y_bar_st

Var_st = (1/(N^2)) * (((710^2)*(1 - (3/710))*((StratifiedSampleFinal$sd_Age_Sample[1]^2) / 3)) + ((9474^2)*(1 - (44/9474)) * ((StratifiedSampleFinal$sd_Age_Sample[2]^2) / 44)))
Var_st

Bound = 2 * (sqrt(Var_st))
Bound

confInt = c(y_bar_st - Bound, y_bar_st + Bound )
confInt
# y_bar_st = (1/10184) * ((710 * mean ) + (9474 * mean))
# y_bar_st
# 
# Var_st = (1/(10184^2)) * (((710^2)*(1 - (3/710))*((sd^2) / 3)) + ((9474^2)*(1 - (44/9474)) * ((sd^2) / 44)))
# Var_st

```

d. yes the true value of the mean, 34.69089, falls within our confidence interval. 
e. Yes we achieved our bound. We chose a bound of 3 and the bound that we got was 3.201517 (This will change every time we run the code so we can fix it once we run all of the code)

5.
For the second sampling method we will be using Simple Random Sampling(SRS)
```{r}
#a.
# From part3, n = 47
n = 47
SRS_sample <- sample(MountEverestData, 47, replace = F)
SRS_sample
```

```{r}
#b.
ggplot(
  data = SRS_sample, 
  aes(x = Sex, y = Age)) +
  geom_boxplot()  + 
  labs(title = "SRS Sex vs. Age")

ggplot(
  data = SRS_sample, 
  aes(x = YearOnly, y = Age)) + 
  geom_point()  + labs(title = "")
```

```{r}
#c.
Finite_Population_correction <- 1-n/N
Finite_Population_correction
#The correction factor in this case can be ignored since 0.9953884 > 0.95

SRSFinal <- 
  SRS_sample %>%
  summarise(n(), sd(Age), mean(Age)) %>%
  rename(count = "n()", 
         sd = "sd(Age)", 
         mean = "mean(Age)")
SRSFinal$s2 <- SRSFinal$sd^2
SRSFinal$V_ybar <- SRSFinal$s2/n
SRSFinal

B1 <- 2*sqrt(Finite_Population_correction*(SRSFinal$s2/n)) # with correction factor
B1
B2 <- 2*sqrt(SRSFinal$s2/n) # w/o correction factor
B2

SRS_confInt = c(SRSFinal$mean - B2, SRSFinal$mean + B2 )
SRS_confInt
```
- The true mean 34.69089 falls within the ~95% confidence 
interval. 

6.
For the third sampling method, we will be using systematic sampling. No pattern so we don't have to have to do successive differences. 
```{r}
plot(MountEverestData$Age)
```
So n - 47
```{r}
#n = Nsigma^2 / (N -1)D+ sigma ^ 2
sdAge = sd(MountEverestData$Age)
sdAge
n = (10184 * sdAge^2) / (((10184 - 1) * ((3^2) / 4)) + sdAge^2)
n
```

So 1 in 216 systematic sampling 
```{r}
10184 / 47
```
so starting point is ____
```{r}
x <- sample(1:216, 1)
x
```
```{r}
y <- c(x)
for (i in 1:46) {
  x = x + 216 
  y <- append(y, x)
}
y
```

```{r}
MountEverestData$ID <- c(1:10184)
head(MountEverestData)

SystematicData <- 
  MountEverestData %>%
  filter(ID %in% y)
SystematicData

mu = mean(SystematicData$Age)
mu

#Vysy = (1 - n/N)s^2/n
V_sy = (1 - (47/10184)) * ((sdAge^2) / 47)
V_sy
Bound = 2 * sqrt(V_sy)
confint = c(mu - Bound, mu + Bound)
confint
```

d. Yes the true mean falls in the confidence interval 